

## Finding important variables to classify clusters

We base our importances on the "classification value" they give to a random
forest (using the implementation in the `ranger` package)

So lets fit the random forest ...
Here we are adding the `warn.imp.method` to prevent a warning message sent by
`ranger` when most of the variables are correlated with the clustering.

Please reffer to the  of the `importance_pvalues` section in the
[ranger documentation](https://cran.r-project.org/web/packages/ranger/ranger.pdf)
when addressing this issue and for more details.

```{r}
library(sctree)
rang_importances <- ranger_importances.Seurat(
    small_5050_mix,
    cluster = "ALL",
    warn.imp.method = FALSE)

```

By default, we obtain a data frame containing only importances with pvalues
under `0.05`.

```{r}
head(rang_importances)
```


# Seurat Interface

As an analogous function to Seurat's `FindAllMarkers`, we offer
`FindAllMarkers_ranger.Seurat`

```{r}
markers <- FindAllMarkers_ranger.Seurat(
  small_5050_mix,
  warn.imp.method = FALSE)

head(markers)

plot.markers <- do.call(rbind, lapply(split(markers, markers$cluster), head, 3))

FeaturePlot(small_5050_mix, unique(plot.markers$gene))
```

Note how variable importances can be high if a marker is either preferentially
present of preferentially absent. Therefore as a pre-filtering step we
implemented a modified version of seurat's "FindMarkers"

```{r}
library(Seurat)
library(sctree)

head(
    sctree::FindMarkers(
        small_5050_mix,
        features = rownames(small_5050_mix@assays$RNA@data),
        ident.1 = 0, test.use = "RangerDE"))

```

```{r}

markers <- sctree::FindAllMarkers(
        small_5050_mix,
        features = rownames(small_5050_mix@assays$RNA@data),
        test.use = "RangerDE")

# Here we just extract the top 3 markers for each cluster
plot.markers <- do.call(rbind, lapply(split(markers, markers$cluster), head, 3))


plot.markers
```

