

```{r include=FALSE}
library(Seurat)
library(sctree)

markers <- sctree::FindAllMarkers(
        small_5050_mix,
        features = rownames(small_5050_mix@assays$RNA@data),
        test.use = "RangerDE")

# Here we just extract the top 3 markers for each cluster
top_markers <- do.call(rbind, lapply(split(markers, markers$cluster), head, 3))
top_markers <- top_markers$gene
```


## Suggesting a gating strategy for the markers

A general strategy to get separate all clusters

```{r}
top_markers
```


```{r}
tree_fit <- fit_ctree(small_5050_mix,
                      genes_use = top_markers,
                      cluster = "ALL")
```

Visualizing the tree as ... a tree ... we can see how our model is a simple
series of yes/no questions.

If we wanted to classifiy a random cell: in the first `node`, we check if the
expression of that gene is higher or lower than a given value, if it is lower,
we proceed to the left, if not we go right. We keep doing that until we have no
more `branches`. This final node will have a predicted cluster, in this plot we
can also see how pure can we expect this group to be and how many of the cells
in our training set clasify as part of it.

```{r fig.height=5, fig.width=12}
plot(tree_fit)
```

When inspecting the tree_fit, we can see a more detailed text representation
of this tree.

```{r}
print(tree_fit)
```

Sometimes one might think that the proposed strategy is too complicated or not
implementable in the experimental settings, in order to add constrians to the fit
one can give additional arguments that will be passed to
`partykit::ctree_control`, such as `maxdepth = 2` (maximum 2 questions per cell)

```{r fig.height=5, fig.width=12}
tree_fit <- fit_ctree(
    small_5050_mix, genes_use = top_markers,
    cluster = "ALL", maxdepth = 2)
print(tree_fit)
plot(tree_fit)
```

Since not all variables are ultimately used in our classifier, one can acces the
ones that were by using `varimp(tree_fit)`

```{r}
partykit::varimp(tree_fit)
plot_flowstyle(small_5050_mix, names(partykit::varimp(tree_fit)))
```


One can also request the package to suggest a specific strategy only for a given
cluster. This function is not expected to give drastically different results in
datasets with few clusters, but it can definitely come usefull when many clusters
are present and one is interested in a specific one.

```{r}
tree_fit <- fit_ctree(small_5050_mix, genes_use = top_markers, cluster = "0")
print(tree_fit)
```

Sometimes it is useful to visualize directly the subset of cells that will be 
"gated" out in each rule, this can be easily achieved by using our implementation
of `plot_gates`


```{r}
plot_gates(small_5050_mix, tree_fit, "6")
```


We have also implemented a way to export these rules as a `garnett` classifier.
for more detail on how the calssifier is implemented please refer to [the garnett
documentation](https://cole-trapnell-lab.github.io/garnett/docs/#constructing-a-marker-file)

```{r}
as.garnett(tree_fit)
```

